---
title: "The stringr package"
author: "Christina Mädge"
date: "21 10 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
```

```{r}
library(readxl)
RAW_recipes_200 <- read_excel("RAW_recipes_200.xlsx")
recipes <- RAW_recipes_200
View(recipes)
glimpse(recipes)
```
```{r}
name <- recipes %>% select(name)
str(name)
recipes %>% 
  select(name) %>% 
  unlist(use.names = FALSE)

vec_name <- pull(recipes, name)
head(vec_name)
str(vec_name)
is.atomic(vec_name)

glimpse(vec_name)

```

# Detect matches

If you want to detect the presence of a pattern match in a string, you use the str_detect() command.
```{r}
str_detect(vec_name, "pizza")
```

If your dataset was bigger, it might be annoying to look through the whole output. So instead of only detecting if there is a recipe containing the word "pizza", you may want to count the recipes including "pizza" in their names. str_count() helps you to count the number of matches in a string. 
```{r}
str_count(vec_name, "pizza")
```

Contrary to expectations, the output is not a number but ...



Now we want to find the indexes of strings that contain a pattern match by using str_which().
```{r}
str_which(vec_name, "pizza")
```

For locating the positions of pattern matches in a string use str_locate().
```{r}
str_locate(vec_name, "pizza")
```


